Distributional Trees and Forests

### 2017-10-12 pkg/disttree/inst/draft/draft.Rnw
small corrections in introduction
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1657 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: d17cfcbb10c4c02e8f93c96abc010a4969462b5a

All these parametric models work very well as long as only additive effects appear in the considered data. A different type of model that can capture non-additive effects as well are regression trees. Here the basic idea is to split the data set into more homogeneous subgroups based on information provided by the explanatory variables and to fit a model/response to each terminal node. One problem that might appear when applying tree algorithms is that they impose steps and abrupt changes on the model, even in situations where effects are rather smooth, as it can be seen in the top left plot of Figure \ref{Devel_treeforest}. One way how to tackle this problem is to not just consider one tree but an ensemble of trees. This is the idea of random forests. A set of trees is built, each on a slightly different data set which can either be a bootstrap or a subsample of the original data set. Then the results of all trees are combined to get a final model which unlike a tree model c!
 an also deal with smooth effects in a reasonable way.\\ 
In distributional trees and forests we now want to combine all these ideas from different fields of statistical modeling in order to benefit from all their advantages.\\
\\
In the past the combination of parametric data models and algorithmic tree models has gained more and more interest since it allows for a more flexible and clear structure. \\
Algorithms that only fit constants to each final node and therefore only model the mean of the response variable (such as CART, Breiman 1984) might lead to very large trees which complicates analyzing and interpreting. Algorithms such as GUIDE (Loh 2002), CRUISE (Kim and Loh 2001) and LOTUS (Chan and Loh 2004) were among the first to provide trees with parametric models in each terminal node. The MOB algorithm (Zeileis, Hothorn and Hornik 2008), an algorithm for model-based recursive partitioning, extended this idea by fitting parametric models to each node and using the gained information of the fitted model to create further splits in inner nodes. Using this algorithm as a framework to build trees LMs and GLMs can be fitted in the nodes of the trees. However, up to now trees which only model the mean of the response variable in its nodes are still more common and widely used.\\
Random forests (\cite{breiman2001random}) applying the CART algorithm to build the trees and cforest applying the ctree algorithm (\cite{hothorn2015partykit}, \cite{hothorn2006unbiased}) have used this idea of combining an ensemble of trees to get smoother effects of the covariates after combining or averaging over the parameters of the trees.\\
\\
But as mentioned before, until now the above listed development in the field of probabilistic modeling has not yet been fully integrated in the idea of algorithmic tree and forest models. Probabilistic models are rarely used in combination with tree algorithms and more flexible models such as GAMLSSs have not yet been applied as models in the nodes of a tree.\\
\\
Therefore, our main goal is to embed the progress made in the field of probabilistic modeling in the idea of regression trees and forests. In particular, more flexible models such as GAMLSSs should be fitted in the nodes of a tree in order to capture location, scale and shape as well as censoring, tail behavior etc. while non-additive effects or interactions of the explanatory variables can be detected automatically by the splitting algorithm used to build the tree. In that way, no variable selection or predefining of variable interactions is necessary and a whole distribution is specified in each node which leads to a wide range of inference methods that can be applied.\\
These (future) steps are illustrated in Figure \ref{Devel_treeforest}, embedded in the development made in the field of tree and forest algorithms.\\

>>>>

%All these parametric models work very well as long as only additive effects appear in the considered data. A different type of model that can capture non-additive effects as well are regression trees. Here the basic idea is to split the data set into more homogeneous subgroups based on information provided by the explanatory variables and to fit a model/response to each terminal node. One problem that might appear when applying tree algorithms is that they impose steps and abrupt changes on the model, even in situations where effects are rather smooth, as it can be seen in the top left plot of Figure \ref{Devel_treeforest}. One way how to tackle this problem is to not just consider one tree but an ensemble of trees. This is the idea of random forests. A set of trees is built, each on a slightly different data set which can either be a bootstrap or a subsample of the original data set. Then the results of all trees are combined to get a final model which unlike a tree model can also deal with smooth effects in a reasonable way.\\ 
%In distributional trees and forests we now want to combine all these ideas from different fields of statistical modeling in order to benefit from all their advantages.\\
%\\
%In the past the combination of parametric data models and algorithmic tree models has gained more and more interest since it allows for a more flexible and clear structure. \\
%Algorithms that only fit constants to each final node and therefore only model the mean of the response variable (such as CART, Breiman 1984) might lead to very large trees which complicates analyzing and interpreting. Algorithms such as GUIDE (Loh 2002), CRUISE (Kim and Loh 2001) and LOTUS (Chan and Loh 2004) were among the first to provide trees with parametric models in each terminal node. The MOB algorithm (Zeileis, Hothorn and Hornik 2008), an algorithm for model-based recursive partitioning, extended this idea by fitting parametric models to each node and using the gained information of the fitted model to create further splits in inner nodes. Using this algorithm as a framework to build trees LMs and GLMs can be fitted in the nodes of the trees. However, up to now trees which only model the mean of the response variable in its nodes are still more common and widely used.\\
%Random forests (\cite{breiman2001random}) applying the CART algorithm to build the trees and cforest applying the ctree algorithm (\cite{hothorn2015partykit}, \cite{hothorn2006unbiased}) have used this idea of combining an ensemble of trees to get smoother effects of the covariates after combining or averaging over the parameters of the trees.\\
%\\
%But as mentioned before, until now the above listed development in the field of probabilistic modeling has not yet been fully integrated in the idea of algorithmic tree and forest models. Probabilistic models are rarely used in combination with tree algorithms and more flexible models such as GAMLSSs have not yet been applied as models in the nodes of a tree.\\
%\\
%Therefore, our main goal is to embed the progress made in the field of probabilistic modeling in the idea of regression trees and forests. In particular, more flexible models such as GAMLSSs should be fitted in the nodes of a tree in order to capture location, scale and shape as well as censoring, tail behavior etc. while non-additive effects or interactions of the explanatory variables can be detected automatically by the splitting algorithm used to build the tree. In that way, no variable selection or predefining of variable interactions is necessary and a whole distribution is specified in each node which leads to a wide range of inference methods that can be applied.\\
%These (future) steps are illustrated in Figure \ref{Devel_treeforest}, embedded in the development made in the field of tree and forest algorithms.\\
